Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-27 introduction.tex
Changed title. Edited intro.

% Recent studies proposed programmable parallel hardware as a way to achieve both
% the performance advantages of hardware and the flexibility of software
% \cite{lba-isca08, flexcore-micro10, harmoni-dsn12}.  While more
% efficient than software, these programmable hardware monitors can still show
% significant performance overheads.  For example,
% Figure~\ref{fig:intro.full_mon} shows the execution time for three monitoring
% techniques normalized to the execution time with no monitoring for an
% FPGA-based parallel monitor. The program being monitored is run on an in-order
% embedded core running at 500 MHz while the FPGA-based monitor runs at 250 MHz.
% We can see that, depending on the monitoring scheme, overheads can easily be
% several tens of percent.  A faster or more complex (e.g., superscalar or
% out-of-order) core is expected to show even more significant overheads.

% % Run-time monitoring overview
% \begin{figure}
%   \begin{center}
%     \vspace{-0.2in}
%     \includegraphics[width=\columnwidth]{figs/full_mon.pdf}
%     \vspace{-0.5in}
%     \caption{Performance overheads of FPGA-based monitors.}
%     \vspace{-0.2in}
%     \label{fig:intro.full_mon} 
%   \end{center}
% \end{figure}

Recent studies have shown that filtering out monitoring checks on clean
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out both
propagation and checking of clean metadata. By keeping a separate set of flags
to track clean metadata, we are able to handle the common case of clean
metadata without the need to read full metadata. By propagating these flags, we
are also able to handle propagation of clean metadata efficiently in hardware.
% By using reduce the
% overheads of monitoringthe dataflow engine
% to track flows of
% initialized metadata, the system is able to filter out events that do not
% correspond to valid metadata, greatly reducing the amount of monitoring that
% must be handled in software. 
Thus, we are able to have the flexibility of software-based monitors with
greatly reduced overheads.

With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. Although full monitoring coverage is always desired,
if these overheads are considered too high by the designer or user then
monitoring must be disabled.
% % Current schemes are all-or-nothing: suffer full performance impact for full
% % coverage.
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  In this paper, we present an architecture that enables a
% trade-off between the overheads and the coverage of a monitoring technique by
% dynamically adjusting the number of monitoring operations.  This system allows
% monitoring to still be performed even when the overhead budget is not
% sufficient for full monitoring.
% Use cases
Thus, we present an architecture that enables a trade-off between overheads and
the monitoring coverage achieved. This is achieved by dynamically dropping
monitoring operations when overheads are too high. We are able to prevent false
positives by extending the dataflow engine to also track these dropped
monitoring flows. We also present three policies on when dropping decisions are
made that cover a trade-off space between ability to match a target overhead
and the monitoring coverage achieved.

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can still be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

The main contributions of this paper are as follows.
\begin{itemize}
  \item Using a dataflow tracking engine to filter out monitoring events in order to reduce overheads.
  \item Extending the dataflow tracking engine to enable monitoring with adjustable overheads
  \item Exploration of the design space and trade-offs in dropping policies for performing partial monitoring
\end{itemize}

@@

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:filter} presents our architecture that use hardware-based DIFT
in order to reduce monitoring overheads. In Section~\ref{sec:drop} we present
how the DIFT engine can be extended to enable adjustable overheads.
Section~\ref{sec:extensions} covers example implementations of three monitoring
schemes. Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

>>>>

% Filtering
Recent studies have shown that filtering out monitoring checks on clean
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out both
propagation and checking of clean metadata. By keeping a separate set of flags
to track clean metadata, we are able to handle the common case of clean
metadata without the need to read full metadata. By propagating these flags, we
are also able to handle propagation of clean metadata efficiently in hardware.
Thus, we are able to have the flexibility of software-based monitors with
greatly reduced overheads.

% Adjustable overhead
With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. 
Traditionally, these overheads have been an all-or-nothing cost. If the
designer finds that the overheads exceed their desired budget, then monitoring
cannot be enabled.  
Thus, we present an architecture that enables a trade-off between overheads and
the monitoring coverage achieved. This is achieved by dynamically dropping
monitoring operations when overheads are too high. We prevent false
positives by extending the dataflow engine to also track these dropped
monitoring flows. We investigate three different policies on when dropping decisions are
made that cover a trade-off space between the ability to match a target overhead
and the monitoring coverage achieved.

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

Thus, the main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
  \item We enable a trade-off between monitoring coverage and overheads by extending the dataflow tracking engine to keep track of dropped monitoring flows.
  \item We explore the design space and trade-offs in dropping policies for performing partial monitoring.
\end{itemize}

@@

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:filter} presents our architecture that uses dataflow-guided filtering
in order to reduce monitoring overheads. In Section~\ref{sec:drop} we present
how the dataflow tracking engine can be extended to enable adjustable overheads.
Section~\ref{sec:extensions} covers example implementations of three monitoring
schemes. Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.
